String Constraints with Regex-Counting and String-Length Solved Efficiently

### 2023-10-19 2023-1-19/sections/algorithms.tex
remove denghang comments

\medskip
\noindent
\emph{Step 1.\denghang{(Intersection)}} For each $i \in [n]$, compute a CEFA $\cB_{i} = \aut_{i,1} \cap \cdots \cap \aut_{i, l_i}$, and let $\Lambda'_{x_i} := \{\cB_i\}$. \qed

@@

Since
$\sum \limits_{\myvec{v} \in \costset(\cB_i)} \eta(\myvec{v}) \myvec{v} = \sum \limits_{\myvec{v} \in \costset(\cB_i) \setminus \{\myvec{0}\}} \eta(\myvec{v}) \myvec{v}$ for each $i \in [n]$ and $\eta \in \parikh(\cU_{\cB_i})$, it turns out that 
the $\myvec{0}$-labeled transitions in $\cU_{\cB_i}$ do not contribute to the final output $\sum \limits_{\myvec{v} \in \costset(\cB_i)} \eta(\myvec{v}) \myvec{v}$. \sout{Therefore, we propose the following size-reduction technique for the NFA $\cU_{\cB_i}$'s.} \denghang{Regardless of alphabet, we take register updates as new alphabet symbols and $\myvec{0}$ as $\epsilon$ symbol. Then we are capable of using standard minimization to reduce the automata size.}

\medskip
\noindent
\emph{Step 2.\denghang{(Automata size reduction)}} For each $i \in [n]$, we view the transitions $(q, \myvec{0}, q')$ in $\cU_{\cB_i}$ as $\varepsilon$-transitions $(q, \varepsilon, q')$, and remove the $\varepsilon$-transitions from $\cU_{\cB_i}$. Then we determinize and minimize the resulting NFA.   \qed

@@

\medskip
\noindent
\emph{Step 3.\denghang{(Parikh image)}} For each $i \in [n]$, we compute an existential LIA formula $\psi_{\cC_i}(\fZ_{\costset_{\cC_i}})$ from $\cC_i$ to represent $\parikh(\cC_i)$. Then we solve the satisfiability of the following formula,
%
\[
\begin{array}{l}
\bigwedge \limits_{i \in [n]} \left(\psi_{\cC_i}(\fZ_{\costset_{\cC_i}}) \wedge \alpha_i\left[\sum\limits_{\myvec{v} \in \costset_{\cC_i}} \anivarz_{i, \myvec{v}} \myvec{v} \big{/}R_{\cB_i}  \right] \right) \ \wedge \\
\ \ \varphi \left[ \sum\limits_{\myvec{v} \in \costset_{\cC_1}} \anivarz_{1, \myvec{v}} \myvec{v} \big{/}R_{\cB_1}, \cdots, \sum\limits_{\myvec{v} \in \costset_{\cC_n}} \anivarz_{n, \myvec{v}} \myvec{v} \big{/}R_{\cB_n} \right].
\end{array}
\]
\qed

>>>>

\medskip
\noindent
\emph{Step 1.(Intersection)} For each $i \in [n]$, compute a CEFA $\cB_{i} = \aut_{i,1} \cap \cdots \cap \aut_{i, l_i}$, and let $\Lambda'_{x_i} := \{\cB_i\}$. \qed

@@

Since
$\sum \limits_{\myvec{v} \in \costset(\cB_i)} \eta(\myvec{v}) \myvec{v} = \sum \limits_{\myvec{v} \in \costset(\cB_i) \setminus \{\myvec{0}\}} \eta(\myvec{v}) \myvec{v}$ for each $i \in [n]$ and $\eta \in \parikh(\cU_{\cB_i})$, it turns out that 
the $\myvec{0}$-labeled transitions in $\cU_{\cB_i}$ do not contribute to the final output $\sum \limits_{\myvec{v} \in \costset(\cB_i)} \eta(\myvec{v}) \myvec{v}$. Therefore, we propose the following size-reduction technique for the NFA $\cU_{\cB_i}$'s. Regardless of alphabet, we take register updates as new alphabet symbols and $\myvec{0}$ as $\epsilon$ symbol. Then we are capable of using standard minimization to reduce the automata size.

\medskip
\noindent
\emph{Step 2.(Automata size reduction)} For each $i \in [n]$, we view the transitions $(q, \myvec{0}, q')$ in $\cU_{\cB_i}$ as $\varepsilon$-transitions $(q, \varepsilon, q')$, and remove the $\varepsilon$-transitions from $\cU_{\cB_i}$. Then we determinize and minimize the resulting NFA.   \qed

@@

\medskip
\noindent
\emph{Step 3.(Parikh image)} For each $i \in [n]$, we compute an existential LIA formula $\psi_{\cC_i}(\fZ_{\costset_{\cC_i}})$ from $\cC_i$ to represent $\parikh(\cC_i)$. Then we solve the satisfiability of the following formula,
%
\[
\begin{array}{l}
\bigwedge \limits_{i \in [n]} \left(\psi_{\cC_i}(\fZ_{\costset_{\cC_i}}) \wedge \alpha_i\left[\sum\limits_{\myvec{v} \in \costset_{\cC_i}} \anivarz_{i, \myvec{v}} \myvec{v} \big{/}R_{\cB_i}  \right] \right) \ \wedge \\
\ \ \varphi \left[ \sum\limits_{\myvec{v} \in \costset_{\cC_1}} \anivarz_{1, \myvec{v}} \myvec{v} \big{/}R_{\cB_1}, \cdots, \sum\limits_{\myvec{v} \in \costset_{\cC_n}} \anivarz_{n, \myvec{v}} \myvec{v} \big{/}R_{\cB_n} \right].
\end{array}
\]
\qed

### 2023-10-19 2023-1-19/sections/cefa.tex
remove denghang comments

In the sequel, we define three CEFA operations: union, intersection, and concatenation. The following section will use these three operations to solve RECL constraints. \denghang{Note that the union, intersection, and concatenation operations are defined in a slightly more involved manner than register automata~\cite{ra} and counter automata~\cite{GGM12}, as a result of the technical choice that the accepting condition is attached to each automaton, not each final state.}

@@

The \emph{union} of $\aut_1$ and $\aut_2$, denoted by $\aut_1 \cup \aut_2$ \sout{is defined in a slightly more involved manner than NFA, as a result of the technical choice that the accepting conditions of CEFA do not distinguish final states}. Two fresh auxiliary registers say $r'_1,r'_2 \not \in R_1 \cup R_2$, are introduced so that the accepting condition knows whether a run is from $\aut_1$ (or $\aut_2$). Specifically, $\aut_1 \cup \aut_2 = (R', Q', \Sigma, \delta', I', F', \alpha')$ where 
\begin{itemize}
\item $R' = R_1 \cup R_2 \cup \{r'_1, r'_2\}$, $Q' = Q_1 \cup Q_2 \cup \{q'_0\}$ with $q'_0 \not \in Q_1 \cup Q_2$, $I' = \{q'_0\}$, 
%
\item $\delta'$ is the union of four transitions sets:
\begin{itemize}
  \item $\{(q'_0, a, q'_1, (\myvec{v_1}, \myvec{0}, 1,0)) \mid \exists q_1 \in I_1.\ (q_1, a, q'_1, \myvec{v_1}) \in \delta_1\}$
  \item $\{(q'_0, a, q'_2, (\myvec{0}, \myvec{v_2}, 0, 1)) \mid \exists q_2 \in I_2.\ (q_2, a, q'_2, \myvec{v_2}) \in \delta_2\}$
  \item $\{(q_1, a, q'_1, (\myvec{v_1}, \myvec{0}, 0, 0)) \mid q_1\not\in I_1.\ (q_1, a, q'_1, \myvec{v_1}) \in \delta_1\}$
  \item $\{(q_2, a, q'_2, (\myvec{0}, \myvec{v_2}, 0, 0)) \mid q_2\not\in I_2.\ (q_2, a, q'_2, \myvec{v_2}) \in \delta_2\}$
\end{itemize}
 where $(\myvec{v_1}, \myvec{0}, 1,0)$ is a vector that updates $R_1$ by $\myvec{v_1}$, updates $R_2$ by $\myvec{0}$, and updates $r'_1,r'_2$ by $1,0$ respectively. Similarly for $(\myvec{0}, \myvec{v_2}, 0, 1)$, and so on.

>>>>

In the sequel, we define three CEFA operations: union, intersection, and concatenation. The following section will use these three operations to solve RECL constraints. Note that the union, intersection, and concatenation operations are defined in a slightly more involved manner than register automata~\cite{ra} and counter automata~\cite{GGM12}, as a result of the technical choice that the accepting condition is attached to each automaton, not each final state.

@@

The \emph{union} of $\aut_1$ and $\aut_2$, denoted by $\aut_1 \cup \aut_2$ is defined in a slightly more involved manner than NFA, as a result of the technical choice that the accepting conditions of CEFA do not distinguish final states. Two fresh auxiliary registers say $r'_1,r'_2 \not \in R_1 \cup R_2$, are introduced so that the accepting condition knows whether a run is from $\aut_1$ (or $\aut_2$). Specifically, $\aut_1 \cup \aut_2 = (R', Q', \Sigma, \delta', I', F', \alpha')$ where 
\begin{itemize}
\item $R' = R_1 \cup R_2 \cup \{r'_1, r'_2\}$, $Q' = Q_1 \cup Q_2 \cup \{q'_0\}$ with $q'_0 \not \in Q_1 \cup Q_2$, $I' = \{q'_0\}$, 
%
\item $\delta'$ is the union of four transitions sets:
\begin{itemize}
  \item $\{(q'_0, a, q'_1, (\myvec{v_1}, \myvec{0}, 1,0)) \mid \exists q_1 \in I_1.\ (q_1, a, q'_1, \myvec{v_1}) \in \delta_1\}$
  \item $\{(q'_0, a, q'_2, (\myvec{0}, \myvec{v_2}, 0, 1)) \mid \exists q_2 \in I_2.\ (q_2, a, q'_2, \myvec{v_2}) \in \delta_2\}$
  \item $\{(q_1, a, q'_1, (\myvec{v_1}, \myvec{0}, 0, 0)) \mid q_1\not\in I_1.\ (q_1, a, q'_1, \myvec{v_1}) \in \delta_1\}$
  \item $\{(q_2, a, q'_2, (\myvec{0}, \myvec{v_2}, 0, 0)) \mid q_2\not\in I_2.\ (q_2, a, q'_2, \myvec{v_2}) \in \delta_2\}$
\end{itemize}
 where $(\myvec{v_1}, \myvec{0}, 1,0)$ is a vector that updates $R_1$ by $\myvec{v_1}$, updates $R_2$ by $\myvec{0}$, and updates $r'_1,r'_2$ by $1,0$ respectively. Similarly for $(\myvec{0}, \myvec{v_2}, 0, 1)$, and so on.

### 2023-10-19 2023-1-19/sections/implementation.tex
remove vspace
remove denghang comments

%!TEX root = ../main.tex
In this section, we list our experimental results. We first present the overall experiment results on RegCoL and AutomatArk benchmark suites. Then we present experiment results of the large-counting part and evaluate the simplification techniques. \denghang{The goal of our experiments is to answer the following questions:
\begin{itemize}
  \item [Q1:] Does the decision procedure perform well~(Sec \ref{subsec:overall_eval}), especially when the counting bounds are large?~(Sec \ref{subsec:large_bounds_eval})
  \item [Q2:] Are the size-reduction techniques practical?~(Sec \ref{subsec:size_reduction_eval})
  \item [Q3:] How effective is the algorithm for solving the nonemptiness problem compared to the case that uses the nuXmv model checker~\cite{atva2020}?~(Sec \ref{subsec:size_reduction_eval})
\end{itemize}
}

\vspace{-2mm}
\subsection{Benchmark Suites and Experiment Setup}\label{sec:bench}
\vspace{-1mm}

Our experiments utilize two benchmark suites, namely, \emph{RegCoL} and \emph{AutomatArk}. \denghang{Other industrial benchmarks suites are not utilized because they contains no counting operators.} There are 48,843 instances in total, and all benchmark instances are in the SMTLIB2 format.
Moreover, it turns out that only 5\% of regexes among the 48,843 instances are non-register-representable (see Section~\ref{subsec:regex2cefa}).

\medskip
\noindent
\emph{RegCoL benchmark suite.} There are 40,628 RECL instances in the RegCoL suite. These instances are generated by extracting regexes with counting operators from the open source regex library \cite{regex_lingua_franca,redos_lenka} and manually constructing a RECL constraint $x \in e \wedge x \in e_{sani} \wedge |x| > 10$ for each regex $e$,
where $e_{sani} \equiv \overline{\Sigma^*(<+ >+'+''+\&)\Sigma^*}$ is a regular expression that sanitizes all occurrence of special characters $<$, $>$, $'$, $''$, or $\&$. 
The expression $e_{sani}$ is introduced in view of the fact that these characters are usually sanitized in Web browsers to alleviate the XSS attacks \cite{malware_detection_3_kudzu,CCH_18}.
\vspace{-0.1mm}

@@

\medskip
\noindent
\denghang{
\emph{Large-Bounds part.} This benchmark suite is composed of 1,969 problem instances with large counting bounds (greater than or equal to $50$) from RegCoL and AutomatArk benchmark suites.  
Moreover, in order to test the performance of the solvers on string constraints with large length bounds as well, we increase the length bound to $200$, that is, $|x| > 200$.
}

@@

\vspace{-2mm}
\subsection{Overall Evaluation}\label{subsec:overall_eval}
\vspace{-1mm}

We evaluate the performance of $\ostrichrecl$ against the state-of-the-art string constraint solvers, including CVC5
\cite{cvc5}, Z3seq \cite{z3seq}, Z3str3
\cite{z3str3}, Z3str3RE \cite{z3str3re}, and OSTRICH
\cite{ostrich2023}, on RegCoL and AutomatArk benchmark suites.
The experiment results can be found in Table~\ref{tab:results_regcol}. Note that we take the result of CVC5 as the ground truth\footnote{Initially,  we used the majority vote of the results of the solvers as the ground truth. Nevertheless, on some problem instances, all the results of the three solvers in the Z3 family are wrong (after manual inspection), thus failing this approach on these instances.}, \denghang{and the results distinct to the ground truth are called \emph{soundness errors}.} We can see that $\ostrichrecl$ solves almost all 48,843 instances, except 182 of them, that is, it solves \textbf{48,662} instances correctly. The number is \textbf{3,908/1,111/12,838/2,306/2,396 more} than the number of instances solved by CVC5/Z3str3RE/Z3str3/Z3seq/OSTRICH respectively.
%
Moreover, $\ostrichrecl$ is the second fastest solver, whose average time on each instance is close to the fastest solver Z3str3RE (\textbf{1.93s}/1.62s). \denghang{To answer \textbf{Q1}, \textbf{$\ostrichrecl$ is the second fastest solver on these benchmark suites, and it solves the most number of instances}.}

%
\begin{table}
  \vspace{-4mm}
  \centering
  \begin{subtable}{0.513\textwidth}
      \centering
      \import{tables}{table_regcol.tex}
      \caption{Overall evaluation, with timeout = 60 seconds.}
      \label{tab:results_regcol}
  \end{subtable}
  \begin{subtable}{0.477\textwidth}
      \centering
      \import{tables}{table_simp.tex}
      \caption{Evaluation of the size-reduction techniques, with timeout = 60 seconds.}
      \label{tab:results_simp}
  \end{subtable}
  \vspace{-2mm}
  \caption{Overall evaluation and the evaluation of the size-reduction techniques}
  \vspace{-8mm}
\end{table}

\subsection{Evaluation on problem instances with large bounds}\label{subsec:large_bounds_eval}
%
\sout{We extract the 1,969 problem instances with large counting bounds (greater than or equal to $50$) from the RegCoL and AutomatArk benchmark suites.  
Moreover, in order to test the performance of the solvers on string constraints with large length bounds as well, we increase the length bound to $200$, that is, $|x| > 200$.}
%
\begin{figure}[ht]
  \centering
  \begin{subfigure}[t]{0.49\textwidth}
    \centering\vskip 0pt
    \includegraphics[width=1\textwidth]{counting_distribution.png}  
    \caption{Distribution of problem instances w.r.t. counting bounds}  
    \label{fig:count_distri}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \centering\vskip 0pt
    \import{tables}{table_large_count.tex}
    \vspace{4.5mm}
    \caption{Experiment results on Large-Bounds part, with timeout = 60 seconds.}
    \label{fig:table_large_count}
  \end{subfigure}
  \vspace{-2mm}
  \caption{Distribution of counting bounds and experiment results for large bounds}
\vspace{-5mm}
\end{figure}

We evaluate the performance of $\ostrichrecl$ on 1,969 instances of the Large-Bounds part. 
The experiment results are put in Fig~\ref{fig:table_large_count}. From the results, $\ostrichrecl$ solves \textbf{1,873} instances correctly, which is \textbf{947/278/563/637/523 more} than that solved by CVC5/Z3str3RE/Z3str3/Z3seq/OSTRICH respectively. Moreover, $\ostrichrecl$ is \textbf{ 6.79/2.88/2.61/5.27/3.95} times faster than CVC5/ Z3str3RE/Z3str3/Z3seq/OSTRICH respectively. \denghang{To answer \textbf{Q1}, we can see that $\ostrichrecl$ \textbf{is far more efficient than the other solvers on problem instances with large bounds.}}

\subsection{Evaluation of the effectiveness of size-reduction techniques}\label{subsec:size_reduction_eval}
We also do experiments to evaluate the effectiveness of the automata size-reduction techniques, that is, Step 2 in Section~\ref{subsec:cefadec}. 
Let us use $\ostrichrecl_{\rm -SIMP}$ to denote $\ostrichrecl$ with the size-reduction techniques removed. 
Meanwhile, we compare $\ostrichrecl$ and $\ostrichrecl_{\rm -SIMP}$ against $\ostrichrecl_{\rm NUXMV}$, where the $\cefadec$ problem is solved by the \textsc{nuXmv} model checker \cite{nuxmv}, instead of the procedure in Section~\ref{subsec:cefadec}. 
The experiment results are put in Table~\ref{tab:results_simp}. 
From the results, $\ostrichrecl$ solves \textbf{1,503 more} instances and is \textbf{2.21} times faster than $\ostrichrecl_{\rm -SIMP}$. \denghang{To answer \textbf{Q2}, \textbf{the size-reduction techniques indeed play a vital role in the performance improvement}}. Moreover, $\ostrichrecl$ solves \textbf{1,798 more} instances and is \textbf{3.13} times faster than $\ostrichrecl_{\rm NUXMV}$. \denghang{To answer \textbf{Q3}, \textbf{our algorithm for solving the nonemptiness problem is much better than the \textsc{nuXmv}-based techniques}}.

>>>>

%!TEX root = ../main.tex
In this section, we list our experimental results. We first present the overall experiment results on RegCoL and AutomatArk benchmark suites. Then we present experiment results of the large-counting part and evaluate the simplification techniques. The goal of our experiments is to answer the following questions:
\begin{itemize}
  \item [Q1:] Does the decision procedure perform well~(Sec \ref{subsec:overall_eval}), especially when the counting bounds are large?~(Sec \ref{subsec:large_bounds_eval})
  \item [Q2:] Are the size-reduction techniques practical?~(Sec \ref{subsec:size_reduction_eval})
  \item [Q3:] How effective is the algorithm for solving the nonemptiness problem compared to the case that uses the nuXmv model checker~\cite{atva2020}?~(Sec \ref{subsec:size_reduction_eval})
\end{itemize}

\subsection{Benchmark Suites and Experiment Setup}\label{sec:bench}

Our experiments utilize two benchmark suites, namely, \emph{RegCoL} and \emph{AutomatArk}. Other industrial benchmarks suites are not utilized because they contains no counting operators. There are 48,843 instances in total, and all benchmark instances are in the SMTLIB2 format.
Moreover, it turns out that only 5\% of regexes among the 48,843 instances are non-register-representable (see Section~\ref{subsec:regex2cefa}).

\medskip
\noindent
\emph{RegCoL benchmark suite.} There are 40,628 RECL instances in the RegCoL suite. These instances are generated by extracting regexes with counting operators from the open source regex library \cite{regex_lingua_franca,redos_lenka} and manually constructing a RECL constraint $x \in e \wedge x \in e_{sani} \wedge |x| > 10$ for each regex $e$,
where $e_{sani} \equiv \overline{\Sigma^*(<+ >+'+''+\&)\Sigma^*}$ is a regular expression that sanitizes all occurrence of special characters $<$, $>$, $'$, $''$, or $\&$. 
The expression $e_{sani}$ is introduced in view of the fact that these characters are usually sanitized in Web browsers to alleviate the XSS attacks \cite{malware_detection_3_kudzu,CCH_18}.

@@

\medskip
\noindent

\emph{Large-Bounds part.} This benchmark suite is composed of 1,969 problem instances with large counting bounds (greater than or equal to $50$) from RegCoL and AutomatArk benchmark suites.  
Moreover, in order to test the performance of the solvers on string constraints with large length bounds as well, we increase the length bound to $200$, that is, $|x| > 200$.

@@

\subsection{Overall Evaluation}\label{subsec:overall_eval}

We evaluate the performance of $\ostrichrecl$ against the state-of-the-art string constraint solvers, including CVC5
\cite{cvc5}, Z3seq \cite{z3seq}, Z3str3
\cite{z3str3}, Z3str3RE \cite{z3str3re}, and OSTRICH
\cite{ostrich2023}, on RegCoL and AutomatArk benchmark suites.
The experiment results can be found in Table~\ref{tab:results_regcol}. Note that we take the result of CVC5 as the ground truth\footnote{Initially,  we used the majority vote of the results of the solvers as the ground truth. Nevertheless, on some problem instances, all the results of the three solvers in the Z3 family are wrong (after manual inspection), thus failing this approach on these instances.}, and the results distinct to the ground truth are called \emph{soundness errors}.} We can see that $\ostrichrecl$ solves almost all 48,843 instances, except 182 of them, that is, it solves \textbf{48,662} instances correctly. The number is \textbf{3,908/1,111/12,838/2,306/2,396 more} than the number of instances solved by CVC5/Z3str3RE/Z3str3/Z3seq/OSTRICH respectively.
%
Moreover, $\ostrichrecl$ is the second fastest solver, whose average time on each instance is close to the fastest solver Z3str3RE (\textbf{1.93s}/1.62s). To answer \textbf{Q1}, \textbf{$\ostrichrecl$ is the second fastest solver on these benchmark suites, and it solves the most number of instances}.

%
\begin{table}
  \centering
  \begin{subtable}{0.513\textwidth}
      \centering
      \import{tables}{table_regcol.tex}
      \caption{Overall evaluation, with timeout = 60 seconds.}
      \label{tab:results_regcol}
  \end{subtable}
  \begin{subtable}{0.477\textwidth}
      \centering
      \import{tables}{table_simp.tex}
      \caption{Evaluation of the size-reduction techniques, with timeout = 60 seconds.}
      \label{tab:results_simp}
  \end{subtable}
  \caption{Overall evaluation and the evaluation of the size-reduction techniques}
\end{table}

\subsection{Evaluation on problem instances with large bounds}\label{subsec:large_bounds_eval}
%
We extract the 1,969 problem instances with large counting bounds (greater than or equal to $50$) from the RegCoL and AutomatArk benchmark suites.  
Moreover, in order to test the performance of the solvers on string constraints with large length bounds as well, we increase the length bound to $200$, that is, $|x| > 200$.
%
\begin{figure}[ht]
  \centering
  \begin{subfigure}[t]{0.49\textwidth}
    \centering\vskip 0pt
    \includegraphics[width=1\textwidth]{counting_distribution.png}  
    \caption{Distribution of problem instances w.r.t. counting bounds}  
    \label{fig:count_distri}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \centering\vskip 0pt
    \import{tables}{table_large_count.tex}
    \caption{Experiment results on Large-Bounds part, with timeout = 60 seconds.}
    \label{fig:table_large_count}
  \end{subfigure}
  \caption{Distribution of counting bounds and experiment results for large bounds}
\end{figure}

We evaluate the performance of $\ostrichrecl$ on 1,969 instances of the Large-Bounds part. 
The experiment results are put in Fig~\ref{fig:table_large_count}. From the results, $\ostrichrecl$ solves \textbf{1,873} instances correctly, which is \textbf{947/278/563/637/523 more} than that solved by CVC5/Z3str3RE/Z3str3/Z3seq/OSTRICH respectively. Moreover, $\ostrichrecl$ is \textbf{ 6.79/2.88/2.61/5.27/3.95} times faster than CVC5/ Z3str3RE/Z3str3/Z3seq/OSTRICH respectively. To answer \textbf{Q1}, we can see that $\ostrichrecl$ \textbf{is far more efficient than the other solvers on problem instances with large bounds.}

\subsection{Evaluation of the effectiveness of size-reduction techniques}\label{subsec:size_reduction_eval}
We also do experiments to evaluate the effectiveness of the automata size-reduction techniques, that is, Step 2 in Section~\ref{subsec:cefadec}. 
Let us use $\ostrichrecl_{\rm -SIMP}$ to denote $\ostrichrecl$ with the size-reduction techniques removed. 
Meanwhile, we compare $\ostrichrecl$ and $\ostrichrecl_{\rm -SIMP}$ against $\ostrichrecl_{\rm NUXMV}$, where the $\cefadec$ problem is solved by the \textsc{nuXmv} model checker \cite{nuxmv}, instead of the procedure in Section~\ref{subsec:cefadec}. 
The experiment results are put in Table~\ref{tab:results_simp}. 
From the results, $\ostrichrecl$ solves \textbf{1,503 more} instances and is \textbf{2.21} times faster than $\ostrichrecl_{\rm -SIMP}$. To answer \textbf{Q2}, \textbf{the size-reduction techniques indeed play a vital role in the performance improvement}. Moreover, $\ostrichrecl$ solves \textbf{1,798 more} instances and is \textbf{3.13} times faster than $\ostrichrecl_{\rm NUXMV}$. To answer \textbf{Q3}, \textbf{our algorithm for solving the nonemptiness problem is much better than the \textsc{nuXmv}-based techniques}.

### 2023-10-19 2023-1-19/sections/introduction.tex
remove vspace
remove denghang comments

\medskip
\noindent 
\emph{Contribution.} In this work, we focus on the class of string constraints with regular membership and string-length function, where the counting operators may occur (called RECL for brevity). We make the following contributions.
\vspace{-0.5mm}
\begin{itemize}
  \item We propose an automata-theoretical approach for solving RECL constraints. 
  Our main idea is to represent the counting operators by cost registers in cost-enriched finite automata (CEFA, see Section~\ref{sec:automaton} for the definition), instead of unfolding them explicitly. The string-length function is modeled by cost registers as well. The satisfiability of RECL constraints is reduced to the nonemptiness problem of CEFA w.r.t. a linear integer arithmetic (LIA) formula. According to the results from~\cite{atva2020}, an LIA formula can be computed to represent the potential values of registers in CEFA.
Thus, the nonemptiness of CEFA w.r.t. LIA formulas can be reduced to the satisfiability of LIA formulas, which is then tackled by off-the-shelf SMT solvers.
  \item We propose techniques to reduce the sizes (i.e. the number of states and transitions) of CEFA, in order to achieve better performance.
  \item Combined with the size-reduction techniques mentioned above, the register representation of regex-counting and string-length in CEFA entails an efficient algorithm for solving RECL constraints. We implement the algorithm on top of OSTRICH, resulting in a string solver called $\ostrichrecl$. 
 %
  \item Finally, we utilize a collection of benchmark suites comprising 48,843 instances in total to evaluate the performance of $\ostrichrecl$. The experiment results show that $\ostrichrecl$ solves the RECL constraints more efficiently than the state-of-the-art string solvers, especially when the counting and length bounds are large (see Table~\ref{tab:results_regcol} and Figure~\ref{fig:table_large_count}). For instance, on 1,873 benchmark instances where the counting bounds are greater than or equal to $50$ and the string lengths are required to be beyond $200$, $\ostrichrecl$ solves at least $278$ more instances than the other solvers, while spending only half or less time per instance on average. 
\end{itemize}
\vspace{-0.5mm}

\medskip
\noindent
\emph{Related work.} 
We discuss more related work about regexes with counting in programming languages.  
The determinism of regexes with counting was investigated in~\cite{GGM12,CL15}. Counting-set automata were proposed in~\cite{redos_lenka,HS+23} to quickly match a subclass of regexes with counting. Moreover, a variant of nondeterministic counter automata, called bit vector automata, was proposed recently in \cite{GKM23} to enable fast matching of a more expressive class of regexes with counting.   Nevertheless, the nonemptiness problem of these automata was not considered, and it is unclear whether these automata models can be used for solving string constraints with regex-counting and string-length. \denghang{For the automata model, the definition of registers of CEFAs resembles the counters of the counter automata class~\cite{GGM12,redos_lenka,HS+23,GKM23}, but CEFAs contain no guards in each transition. Register automata~\cite{ra} and symbolic register automata~\cite{sra} utilize registers to store input variables (i.e., the letters of the alphabet), on which only equality and inequality are permitted. In contrast, CEFAs utilize registers to store integers constrained by linear arithmetic formulae.}
Real-world regexes in programming languages include features beyond classical regular expressions, e.g., the greedy/lazy Kleene star, capturing groups, and back references. Real-world regexes have been addressed in symbolic execution of Javascript programs~\cite{LMK19} and string constraint solving~\cite{CF+22}. Nevertheless, the counting operators are still unfolded explicitly in~\cite{CF+22}. \sout{Solving string constraints involving lengths has also been investigated by various researchers, see e.g. \cite{WC+18,atva2020,z3str3re}.} \denghang{Except for the string solvers mentioned before, some related works about string constraints with the length function exist. Trau~\cite{z3trau} relies on a Counter-Example Guided Abstraction Refinement (CEGAR) framework in which the under- and over-approximation of the automata are iteratively refined to solve length and regex constraints. Slent~\cite{WC+18} transfers the string and length constraints to length-encoded automata, and leverages a symbolic model checker as an engine for the emptiness checking of automata. However, both Trau and Slent do not support counting operators.}

>>>>

\medskip
\noindent 
\emph{Contribution.} In this work, we focus on the class of string constraints with regular membership and string-length function, where the counting operators may occur (called RECL for brevity). We make the following contributions.
\begin{itemize}
  \item We propose an automata-theoretical approach for solving RECL constraints. 
  Our main idea is to represent the counting operators by cost registers in cost-enriched finite automata (CEFA, see Section~\ref{sec:automaton} for the definition), instead of unfolding them explicitly. The string-length function is modeled by cost registers as well. The satisfiability of RECL constraints is reduced to the nonemptiness problem of CEFA w.r.t. a linear integer arithmetic (LIA) formula. According to the results from~\cite{atva2020}, an LIA formula can be computed to represent the potential values of registers in CEFA.
Thus, the nonemptiness of CEFA w.r.t. LIA formulas can be reduced to the satisfiability of LIA formulas, which is then tackled by off-the-shelf SMT solvers.
  \item We propose techniques to reduce the sizes (i.e. the number of states and transitions) of CEFA, in order to achieve better performance.
  \item Combined with the size-reduction techniques mentioned above, the register representation of regex-counting and string-length in CEFA entails an efficient algorithm for solving RECL constraints. We implement the algorithm on top of OSTRICH, resulting in a string solver called $\ostrichrecl$. 
 %
  \item Finally, we utilize a collection of benchmark suites comprising 48,843 instances in total to evaluate the performance of $\ostrichrecl$. The experiment results show that $\ostrichrecl$ solves the RECL constraints more efficiently than the state-of-the-art string solvers, especially when the counting and length bounds are large (see Table~\ref{tab:results_regcol} and Figure~\ref{fig:table_large_count}). For instance, on 1,873 benchmark instances where the counting bounds are greater than or equal to $50$ and the string lengths are required to be beyond $200$, $\ostrichrecl$ solves at least $278$ more instances than the other solvers, while spending only half or less time per instance on average. 
\end{itemize}

\medskip
\noindent
\emph{Related work.} 
We discuss more related work about regexes with counting in programming languages.  
The determinism of regexes with counting was investigated in~\cite{GGM12,CL15}. Counting-set automata were proposed in~\cite{redos_lenka,HS+23} to quickly match a subclass of regexes with counting. Moreover, a variant of nondeterministic counter automata, called bit vector automata, was proposed recently in \cite{GKM23} to enable fast matching of a more expressive class of regexes with counting.   Nevertheless, the nonemptiness problem of these automata was not considered, and it is unclear whether these automata models can be used for solving string constraints with regex-counting and string-length. For the automata model, the definition of registers of CEFAs resembles the counters of the counter automata class~\cite{GGM12,redos_lenka,HS+23,GKM23}, but CEFAs contain no guards in each transition. Register automata~\cite{ra} and symbolic register automata~\cite{sra} utilize registers to store input variables (i.e., the letters of the alphabet), on which only equality and inequality are permitted. In contrast, CEFAs utilize registers to store integers constrained by linear arithmetic formulae.
Real-world regexes in programming languages include features beyond classical regular expressions, e.g., the greedy/lazy Kleene star, capturing groups, and back references. Real-world regexes have been addressed in symbolic execution of Javascript programs~\cite{LMK19} and string constraint solving~\cite{CF+22}. Nevertheless, the counting operators are still unfolded explicitly in~\cite{CF+22}. Solving string constraints involving lengths has also been investigated by various researchers, see e.g. \cite{WC+18,atva2020,z3str3re}. Except for the string solvers mentioned before, some related works about string constraints with the length function exist. Trau~\cite{z3trau} relies on a Counter-Example Guided Abstraction Refinement (CEGAR) framework in which the under- and over-approximation of the automata are iteratively refined to solve length and regex constraints. Slent~\cite{WC+18} transfers the string and length constraints to length-encoded automata, and leverages a symbolic model checker as an engine for the emptiness checking of automata. However, both Trau and Slent do not support counting operators.
